\Large{All in the family:\\German twin finishing times in\\the 2016 women's Olympic marathon}

### 2017-02-23 paper/paper.tex
updates to simulations

Figure \ref{fig:scatter} contains two plots, both of which describe
how finishing times from the Rio women's marathon varied as a function
of athletes' personal best times and half marathon split times.  The
points in the plots are colored by twin/triplet status, and both plots
contain second-order polynomial regression lines.  The ticks along the
horizontal axes in both plots indicate times, either personal bests or
half marathon splits, of runners who earned DNF
results.\label{scatterplots}

@@

To estimate the conditional distribution of each Rio final result, we
assume that each runner's marathon time \Yi\ is distributed normally
with a mean that is a quadratic function of the runner's marathon
skill $X_i$ and her age $Z_i$. Hence:
$$\Yi \mid X_i \sim N\left(\beta_0 + \beta_1\,X_{i} + \beta_2\,X^2_{i}  + \beta_3\,Z_{i} + e_i, \sigma\right),$$
%%
where $N\left(\cdot,\cdot\right)$ denotes a normal
distribution. Taking each runner's personal best marathon time as a
proxy for her underlying marathon skill, we estimate $\beta_1$,
$\beta_2$, $\beta_3$ and $\sigma$ using ordinary least squares on all
finishing Rio marathoners.  We exclude the Hahner/Kim twins and Luik
triplets from the sample so that that our estimates are not affected
by the twin/triplet finishes which, theoretically, could reflect the
results of coordination.  For a simulated marathon, we draw a runner's
time from an estimated distribution and condition on the runner's
personal best marathon time and age.  Once a race is simulated for all
runners, twins and triplets included, we record both the time between
Anna and Lisa Hahner's simulated finishes and the difference in their
simulated ranks.  We then simulate a new race---drawing a new set of
finishing times---and record the same quantities. The steps of the
simulation are as follows.
\begin{enumerate}
\item Ignoring twins and triplets, estimate with least squares a
  linear model that predicts a runner's finishing time $Y_i$ based on
  her pre-Olympic personal best time $X_i$, her best time squared
  $X^2_{i}$, and her age $Z_i$.
\item Extract the resulting coefficient vector, estimated covariance
  matrix, and estimated regression variance from this model.
\item For each simulated rate, draw intercept and slope estimates
  $\tilde{\beta}_0$, $\tilde{\beta}_1$, $\tilde{\beta}_2$, and
  $\tilde{\beta}_3$, respectively, from a multivariate normal
  distribution with mean equal to the previously estimated coefficient
  vector and covariance equal to the previously estimated covariance
  matrix.
\item For each runner, draw an error $\tilde{e}$ from a normal
  distribution with mean zero and a standard deviation equal to the
  standard deviation of the original regression model's residuals.
\item Predict each runner's final result by combining the randomly
  generated beta coefficients and individual error terms,
  $\tilde{y} = \tilde{\beta}_0 + \tilde{\beta}_1\,X_i +
  \tilde{\beta}_2\,X_i^2 + \tilde{\beta}_3\,Z_i + \tilde{e}$.
\item Eliminate each runner from the simulated race with a probability
  equal to the observed fraction of marathoners who did not finish the
  marathon.
% \item Calculate the difference in time and difference in ranking
%   between Anna Hahner and Lisa Hahner, assuming both finished the
%   race.
\item Repeat above steps 10,000 times.
\end{enumerate}

As these steps illustrate, our simulation repeatedly draws random
coefficient vectors, and this captures variability in what we know
about the relationship between runner talent and age and runer finish
times.  In addition, the simulation draws random disturbances for each
runner, conditional on our original estimate of regression variance;
these disturbances capture variability in runner finishing times,
notwithstanding underlying runner talent.  Importantly, the
disturbances that we draw are independent across runners.  This means
that the finishing order among runners for each simulated race will
vary.

>>>>

Figure \ref{fig:scatter} contains two plots, both of which describe
how finishing times from the Rio women's marathon varied as a function
of athletes' personal best times and half marathon split times.  The
points in the plots are colored by twin/triplet status, and both plots
contain %second-order polynomial regression lines. 
a dashed line representing the line of best fit.
The ticks along the
horizontal axes in both plots indicate times, either personal bests or
half marathon splits, of runners who earned DNF
results.\label{scatterplots}

@@

To estimate the conditional distribution of each Rio final result, we
assume that each runner's marathon time \Yi\ is distributed normally
with a mean that is a linear function of the runner's marathon
skill $X_i$ and her age $Z_i$. Hence:
$$\Yi \mid X_i \sim N\left(\beta_0 + \beta_1\,X_{i}  + \beta_2\,Z_{i} + e_i, \sigma\right),$$
%%
where $N\left(\cdot,\cdot\right)$ denotes a normal
distribution. Taking each runner's personal best marathon time as a
proxy for her underlying marathon skill, we estimate $\beta_1$,
$\beta_2$, $\beta_3$ and $\sigma$ using ordinary least squares on all
finishing Rio marathoners.  We exclude the Hahner/Kim twins and Luik
triplets from the sample so that that our estimates are not affected
by the twin/triplet finishes which, theoretically, could reflect the
results of coordination.  For a simulated marathon, we draw a runner's
time from an estimated distribution and condition on the runner's
personal best marathon time and age.  Once a race is simulated for all
runners, twins and triplets included, we record both the time between
Anna and Lisa Hahner's simulated finishes and the difference in their
simulated ranks.  We then simulate a new race---drawing a new set of
finishing times---and record the same quantities. The steps of the
simulation are as follows.
\begin{enumerate}
\item Ignoring twins and triplets, estimate with least squares a
  linear model that predicts a runner's finishing time $Y_i$ based on
  her pre-Olympic personal best time $X_i$ and her age $Z_i$.
\item Extract the resulting coefficient vector, estimated covariance
  matrix, and estimated regression variance from this model.
\item For each simulated rate, draw intercept and slope estimates
  $\tilde{\beta}_0$, $\tilde{\beta}_1$, and
  $\tilde{\beta}_2$, respectively, from a multivariate normal
  distribution with mean equal to the previously estimated coefficient
  vector and covariance equal to the previously estimated covariance
  matrix.
\item For each runner, draw an error $\tilde{e}$ from a normal
  distribution with mean zero and a standard deviation equal to the
  standard deviation of the original regression model's residuals.
\item Predict each runner's final result by combining the randomly
  generated beta coefficients and individual error terms,
  $\tilde{y} = \tilde{\beta}_0 + \tilde{\beta}_1\,X_i +  \tilde{\beta}_2\,Z_i + \tilde{e}$.
\item Eliminate each runner from the simulated race with a probability
  equal to the observed fraction of marathoners who did not finish the
  marathon.
% \item Calculate the difference in time and difference in ranking
%   between Anna Hahner and Lisa Hahner, assuming both finished the
%   race.
\item Repeat above steps 10,000 times.
\end{enumerate}

As these steps illustrate, our simulation repeatedly draws random
coefficient vectors, and this captures variability in what we know
about the relationship between runner talent and age and runner finish
times.  In addition, the simulation draws random disturbances for each
runner, conditional on our original estimate of regression variance;
these disturbances capture variability in runner finishing times,
notwithstanding underlying runner talent.  Importantly, the
disturbances that we draw are independent across runners.  This means
that the finishing order among runners for each simulated race will
vary.
